---
title: "Facial Recognition"
output: 
    github_document:
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tweetrmd)
library(webshot2)
library(default)

# set screenshots to smaller size than default
default(tweet_screenshot) <- list(scale = 1)

```


## Materials

### Materials for Book Club

This session we will be discussing **Facial Recognition** - the good, the bad and the ugly! No need to read everything - pick a few! If you've just got time for one, pick the one in yellow.

**Facial recognition is bad?**

* Facial recognition is the plutonium of AI. by Luke Stark, Published in Association for Computer Machinery it argues that facial recognition has few legitimate uses [article](https://static1.squarespace.com/static/59a34512c534a5fe6721d2b1/t/5cb0bf02eef1a16e422015f8/1555087116086/Facial+Recognition+is+Plutonium+-+Stark.pdf)

**Facial recognition is good?**

* BBC [article](https://www.bbc.co.uk/news/uk-scotland-edinburgh-east-fife-47614890) on pigs - Facial recognition tool &#39;could help boost pigs' wellbeing
* Collateral [article](https://www.collater.al/en/facebook-lim-si-ping-installation/) on art - Art director Lim Si Ping has created Facebook, the first digital book that, based on facial recognition, tells the perfect story for you 
* China uses facial recognition technology to cut down on toilet paper theft [article](https://www.washingtonpost.com/news/morning-mix/wp/2017/03/21/china-uses-facial-recognition-software-to-crack-down-on-toilet-paper-theft/?noredirect=on&amp;utm_term=.cf121bf595a6)

**Should we get ugly? Activism**

For a general background, have a browse of the **Algorithmic Justice League** [website](https://www.ajlunited.org/). The recent resistance to Amazon's use of FR is a case in point. Further reading below:

* **Joy Buolamwini's response to Amazon** (Medium [post](https://medium.com/@Joy.Buolamwini/response-racial-and-gender-bias-in-amazon-rekognition-commercial-ai-system-for-analyzing-faces-a289222eeced)) - _Response: Racial and Gender bias in Amazon Rekognition — Commercial AI System for Analyzing Faces_
* **Letter from 'Concerned researchers'** (Medium [post](https://medium.com/@bu64dcjrytwitb8/on-recent-research-auditing-commercial-facial-analysis-technology-19148bda1832)) - *On Recent Research Auditing Commercial Facial Analysis Technology*
* [**AI Now Letter**](https://ainowinstitute.org/dhcr-amici-letter-043019.pdf) re: Dangers of using facial recognition for housing.
  + Write up of the letter (The Verge [article](https://www.theverge.com/2019/4/3/18291995/amazon-facial-recognition-technology-rekognition-police-ai-researchers-ban-flawed)) - AI researchers tell Amazon to stop selling 'flawed' facial recognition to the police



### Further Reading 
_Reports _

* Oxfam [report](https://policy-practice.oxfam.org.uk/publications/biometrics-in-the-humanitarian-sector-620454) on _Biometrics in the Humanitarian Sector _
* Short report on _police use of facial recognition_ (UK Govt, [here](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/781745/Facial_Recognition_Briefing_BFEG_February_2019.pdf))
* [US Government&#39;s](https://www.ntia.doc.gov/files/ntia/publications/aclu_an_ethical_framework_for_face_recognition.pdf)_Ethical Framework for Facial Recognition_
* SAFE pledge from the Algorithmic Justice League [_https://www.safefacepledge.org/pledge_](https://www.safefacepledge.org/pledge)

_Articles _

* [Article](https://howwegettonext.com/silicon-valley-thinks-everyone-feels-the-same-six-emotions-38354a0ef3d7): _Silicon Valley Thinks Everyone Feels the Same Six Emotions _
* FT [article](https://www.ft.com/content/c2300a60-5a2b-11e9-939a-341f5ada9d40): _Taser stun gun maker files facial recognition patents_
* Non-consensual use of data for facial recognition training (Slate, [here](https://slate.com/technology/2019/03/facial-recognition-nist-verification-testing-data-sets-children-immigrants-consent.html)) - _The Government Is Using the Most Vulnerable People to Test Facial Recognition Software_
* New York Times - _One Month, 500,000 Face Scans: How China Is Using A.I. to Profile a Minority_ ([here](https://www.nytimes.com/2019/04/14/technology/china-surveillance-artificial-intelligence-racial-profiling.html))

## Questions

### Facilitator Prompt Questions

#### Part 1
* Generally right now, ethics of facial recog - biggest concern? Biggest benefit? 
* Should there be bans in any area? (housing, law enforcement, drone targeting)?
* The pig article - use on animals ok? Why or why not? 
* What if DK asked to work on a project for a charity - what process? What about if a local government?
* Ethics of data collection/use/normalisation (scraping from flickr, those google “which art does your face look like” stealth toys and apple’s animoji)?

#### Part 2
* Is inaccuracy the biggest ethical concern/what if it was 100% accurate? (Joy Buolamwini’s work on race, Luke Stark article)
* Adversarial techniques - feasible? Legal? Will there be arms race?
* Where’s it gonna be in 10 years time? Where should it be in 10 years time?



## Outputs

### Live Tweets/Commentary

For tweets from the evening see [here](https://twitter.com/search?q=(%23dkbookclub%20OR%20%23dkethicsbookclub)%20until%3A2019-05-23%20since%3A2019-05-21&src=typed_query&f=live).

```{r} 
tweet_screenshot("https://twitter.com/DataKindUK/status/1131258299490406400")

tweet_screenshot("https://twitter.com/DataKindUK/status/1131259359848275969")

tweet_screenshot("https://twitter.com/DataKindUK/status/1131260753774821381")

tweet_screenshot("https://twitter.com/DataKindUK/status/1131279657414868992")


```

### Blog

[**Facing up to the power of AI**](https://medium.com/datakinduk/facing-up-to-the-power-of-ai-b013c3ecda61)

## Feedback


## Notes or other comments

