---
title: "Self-Driving Cars"
output: 
    github_document:
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tweetrmd)
library(webshot2)
library(default)

# set screenshots to smaller size than default
default(tweet_screenshot) <- list(scale = 1)

```


## Materials

### Materials for Book Club

This session we will be discussing **Self-Driving Cars**.


* [**Car Wars**](https://this.deakin.edu.au/self-improvement/car-wars) **(Cory Doctorow 2016)**
* Here's an alternative shorter take from [Guardian](https://www.theguardian.com/technology/2015/dec/23/the-problem-with-self-driving-cars-who-controls-the-code) by the same author



### Further Reading 
For those who want more....

_A primer_

* And for those who would like it, a primer on self-driving cars: 
**Self-driving car technology explained to non experts** ([Medium](https://medium.com/swlh/everything-about-self-driving-cars-explained-for-non-engineers-f73997dcb60c))

_Impact on inequality _

* [Academic paper](https://arxiv.org/pdf/1902.11097.pdf): Predictive Inequity in Object Detection _- systems unable to identify across skin colors. _
* In this work, we investigate whether state-of-the-art object detection systems have equitable predictive performance on pedestrians with different skin tones. This work is motivated by many recent examples of ML and vision systems displaying higher error rates for certain demographic groups than others. We annotate an existing large scale dataset which contains pedestrians, BDD100K, with Fitzpatrick skin tones in ranges [1-3] or [4-6]. We then provide an in depth comparative analysis of performance between these two skin tone groupings, finding that neither time of day nor occlusion explain this behavior, suggesting_ **this disparity is not merely the result of pedestrians in the 4-6 range appearing in more difficult scenes for detection** _. We investigate to what extent time of day, occlusion, and reweighting the supervised loss during training affect this predictive bias._
* See also news article on this paper, A new study finds a potential risk with self-driving cars: failure to detect dark-skinned pedestrians ([_Vox_](https://www.vox.com/future-perfect/2019/3/5/18251924/self-driving-car-racial-bias-study-autonomous-vehicle-dark-skin)_, March 2019)
* Research paper: Self-Driving Cars: The Impact on People with Disabilities ([Ruderman Foundation](https://rudermanfoundation.org/white_papers/self-driving-cars-the-impact-on-people-with-disabilities/), January 2017)

_Questions of morality…._

* Interactive scenarios: The Moral Machine ([MIT](http://moralmachine.mit.edu/))
  + Academic paper: The Moral Machine Experiment ([Nature](https://www.researchgate.net/publication/328491510_The_Moral_Machine_Experiment), November 2018) - _this article lays some interesting foundations on the absence of universal moral rules (differing by country). 
  + See also the review of the paper, providing some more context:_ Self-driving car dilemmas reveal that moral choices are not universal ([Nature](https://www.nature.com/articles/d41586-018-07135-0#ref-CR1), October 2018)
  + News article: Will your driverless car be willing to kill you to save the lives of others? ([Guardian](https://www.theguardian.com/science/2016/jun/23/will-your-driverless-car-be-willing-to-kill-you-to-save-the-lives-of-others), June 2016)
*People like utilitarian view in theory, but don't want to buy a car programmed that way*

_Industry responses_

* News article: 11 companies propose guiding principles for self-driving vehicles ([Venture Beat](https://venturebeat.com/2019/07/02/self-driving-car-report-safety-first-for-automated-driving/), July 2019)
* Comment piece: Bob Lutz: Kiss the good times goodbye ([Automotive News](https://www.autonews.com/article/20171105/INDUSTRY_REDESIGNED/171109944/bob-lutz-kiss-the-good-times-goodbye), November 2017) - _Bob Lutz (ex-head of General Motors) gives a summary of where he thinks things might be heading _

_Technical challenges and solutions _

* Mobileye CEO Amnon Shashua&#39;s recent lecture at MIT on the challenges of reaching full autonomy ([video](https://cbmm.mit.edu/video/successes-and-challenges-modern-artificial-intelligence))
* The challenge of adversarial attacks ([article](https://www.technologyreview.com/s/613170/emtech-digital-dawn-song-adversarial-machine-learning/))
* Mobileye's 2018 [paper](https://arxiv.org/pdf/1708.06374.pdf) laying out a reasonable model of self-driving behavior

## Questions

### Facilitator Prompt Questions

- Discuss whether Fictional stories have a place to discuss ethical consequences

- Discuss whether you can boil down self-driving cars to the trolley problem? Is this possible, if so why not?

- Reflect on how you felt about the trolley problem, before and after the reading. Discuss if your view has remained the same or changed accordingly.

- Can a manufacturer stop you from changing its programming so that your car can never kill you?

- Discuss whether an owner should hold responsibility for the self-driving car if there is an accident? Who else should take responsibility? Why?

- Would you modify the software in your self-driving car to protect yourself, even if it was against the law?

- Can self-driving and 'dumb' cars co-exist on the roads?

- When buying a product embedded with technology, who owns the product? The code? Why?

- Discuss the following statement:
*"If self-driving cars can only be safe if we are sure no one can reconfigure them without manufacturer approval, then they will never be safe."*


## Outputs

### Live Tweets/Commentary

For tweets from the evening see [here](https://twitter.com/search?q=(%23dkbookclub%20OR%20%23dkethicsbookclub)%20until%3A2019-08-16%20since%3A2019-08-14&src=typed_query&f=live).

```{r} 
tweet_screenshot("https://twitter.com/DataKindUK/status/1161697758778146816")

tweet_screenshot("https://twitter.com/DataKindUK/status/1161703369733476354")

tweet_screenshot("https://twitter.com/DataKindUK/status/1161707870582431747")
```

### Blog


## Feedback


## Notes or other comments
This was our first try of a book club event with one main reading and then various other sources "for interest". Also our first one using fiction! The *Car Wars* piece was quite good as an unusual jumping-off point for the discussion. 

