---
title: "AI & bad credit: the impact of automation on financial inclusion"
output: 
    github_document:
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tweetrmd)
library(webshot2)
library(default)

# set screenshots to smaller size than default
default(tweet_screenshot) <- list(scale = 1)

```


## Materials

### Materials for Book Club

Our next data science ethics bookclub is on _AI and bad credit: the impact of automation on financial inclusion._ **You are welcome to pick from this reading list, depending on your interest and the time you have:**

- **Blog: &#39;** We Didn&#39;t Explain the Black Box – We Replaced it with an Interpretable Model&#39;, about the FICO explainable credit score competition winner [here](https://community.fico.com/s/blog-post/a5Q2E0000001czyUAA/fico1670)
- **Academic article** on fairness in credit risk evaluation, &#39;Context-conscious fairness in using machine learning to make decisions&#39; [here](https://sigai.acm.org/static/aimatters/5-2/AIMatters-5-2-07-Lee.pdf)
- **Government paper** : The Centre for Data Ethics and Innovation look at AI and Personal Insurance [here](https://www.gov.uk/government/publications/cdei-publishes-its-first-series-of-three-snapshot-papers-ethical-issues-in-ai/snapshot-paper-ai-and-personal-insurance)
- **News article** on how companies can use data with low levels of regulation, &#39;The new lending game, post-demonetisation&#39; [here](https://tech.economictimes.indiatimes.com/news/technology/the-new-lending-game-post-demonetisation/56367457)
- **Academic article** on discrimination in consumer lending [here](https://faculty.haas.berkeley.edu/morse/research/papers/discrim.pdf)

### Further Reading

* [https://www.fca.org.uk/publication/research/price\_discrimination\_in\_financial\_services.pdf](https://www.fca.org.uk/publication/research/price_discrimination_in_financial_services.pdf)

* [https://www.fca.org.uk/publication/feedback/fs19-04.pdf](https://www.fca.org.uk/publication/feedback/fs19-04.pdf)


## Questions

### Facilitator Prompt Questions

**Note:** Questions are based on [CDEI meeting] ](https://twitter.com/peterkwells/status/1178597424669089792/photo/1)

**Warm up**

- Which of the papers from the recommended reading list did you read? What were your main takeaways?
- What are the main benefits of data science &amp; AI into the financial sector? Prompts (if conversation is slow:
  - More efficient financial markets
  - Increased access to financial products e.g fin tech innovation, or existing finance services
  - Faster access to financial products
  - Personalised services e.g. AI powered chatbots
  - Detecting vulnerable groups earlier to provide support
  - Better detection of economic crime- fraud/money laundering
  - Increased detection of cyber threats
  - Increase finance services efficiency
  - Better risk assessment therefore low costs for customers
  - Efficient compliance (regulatory technology)
- Who are the beneficiaries? Which group benefits the most

**Negative consequences?**

- What do you think some of the major risks within financial services sector arising from the application of data driven technology and AI? Prompts (if conversation is slow:
  - Cyber attacks
  - Extreme market movement (algorithms used to automate trade decisions) &amp; difficulties monitoring algorithmic trade (black box)
  - Lack of transparency
  - Lack of explainability
  - Loss of public trust in financial systems
  - Use of non traditional finance data e.g social media encroaches on privacy
  - Digital exclusion- what about people who don&#39;t use digital- they don&#39;t create lots of data!
  - Preferential access to people who are willing to give lots of data e.g. provide car/fitness sensors to insurers
  - Bias due to use of historical data to predict forwards
  - Consumer disempowerment- asymmetric power as financial services know more about a customer than the customer itself
  - Regulators unable to keep up with AI due to lack of resources
  - Algorithmic collusion (outside of the financial institution)
  - Increased surveillance of finance workers
  - Deskilling of financial workers/ workers have excessive trust in algorithmic recommendations
  - Data monopolies
  - Fear of potential risks reduces AI takeup which could be a benefit
  - Excessive data retention
  - Increased inequality
  - Non-alignment with societal goals

- How do these risks arise?
- Which groups are most affected and how?
-

**General discussion question**

- Consider yourself/family/friends as a consumer (e.g. for credit, car insurance, health insurance). What data would you share? What would be your concerns about its use? What would you want to know about a company&#39;s decision (explainability etc)? And - how should this be regulated/controlled? How have the reading materials and discussion so far changed how you thought about any of this?

**Governance**

- What governance is or should be in place to mitigate for negative consequences?
  - Legislation and regulation
  - Technical solutions (can lead to discussions around fairness- see Michelle Lee&#39;s article)
  - Soft governance (standards and codes)
  - Anything else?

## Outputs

### Live Tweets/Commentary

For tweets from the evening see [here](https://twitter.com/search?q=(%23dkbookclub%20OR%20%23dkethicsbookclub)%20until%3A2019-10-17%20since%3A2019-10-15&src=typed_query&f=live).

```{r} 

tweet_screenshot("https://twitter.com/DataKindUK/status/1184540754749874180")
tweet_screenshot("https://twitter.com/mcd_ails/status/1184577451709612032")

```

### Blog

[How do we leverage the possibilities of AI for greater financial inclusion?](https://medium.com/datakinduk/how-do-we-leverage-the-new-opportunities-of-ai-for-greater-financial-inclusion-a8a326fd7f3b)

## Feedback


## Notes or other comments


