---
title: "Book Club Topic Template"
output: 
    github_document:
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# allow embedding of tweets to show threads from our bookclubs
library(tweetrmd)
library(webshot2)
library(default)

# set screenshots to smaller size than default
default(tweet_screenshot) <- list(scale = 1)

```

# Topic Name



## Materials

### Materials for Book Club

[consider putting highest priority item in **bold**]

### Further Reading 
This may be a longlist of materials considered, or further reading given as options for attendees.

## Questions

### Pre-Event Prompt Questions

### Facilitator Prompt Questions

## Outputs

### Live Tweets/Commentary

For tweets from the evening see [here]() (Add link as appropriate dated Twitter search eg https://twitter.com/search?q=(%23dkbookclub%20OR%20%23dkethicsbookclub)%20until%3A2019-11-28%20since%3A2019-11-26&src=typed_query&f=live).
```{r echo = TRUE}

## If you have a twitter thread during the book club find representative tweet and embed
## uses the "tweetrmd" https://github.com/gadenbuie/tweetrmd package. 

#----------------------------------------------------------------------
#UNCOMMENT AND REPLACE URL
#tweet_screenshot("https://twitter.com/DataKindUK/status/1113134735243595786")


```

### Blog

## Feedback

## Notes or other comments

